---
title: "LDA"
author: "konda varshith"
date: "12/10/2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---


```{r setup, include=FALSE,echo = FALSE , message=FALSE , error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(MASS)
```

```{r,echo = FALSE , message=FALSE , error=FALSE}


data <- iris
```




### Scatter plot of Sepal length vs Sepal Width

```{r,echo = FALSE , message=FALSE , error=FALSE}
ggplot(data ,aes(x=Sepal.Width,y=Sepal.Length , color  = Species))+
  geom_point()
```



#### Mean

```{r , message=FALSE , error=FALSE  ,echo=FALSE}
setosa <- filter(data ,Species == "setosa")
versicolor <- filter(data ,Species == "versicolor")
virginica <- filter(data ,Species == "virginica")

setosa_mean <-  c(mean(setosa$Sepal.Length) , mean(setosa$Sepal.Width))
versicolor_mean <-  c(mean(versicolor$Sepal.Length) , mean(versicolor$Sepal.Width))
virginica_mean <-  c(mean(virginica$Sepal.Length) , mean(virginica$Sepal.Width))

```

### Covariance 

```{r , message=FALSE , error=FALSE,echo=FALSE}


setosa_covariance <- cov(setosa[1:2])
versicolor_covariance <- cov(versicolor[1:2])
virginica_covariance <- cov(virginica[1:2])
setosa_covariance
virginica_covariance
versicolor_covariance
covariance_matrix <- c(setosa_covariance,virginica_covariance,versicolor_covariance)
covariance_matrix <- as.matrix(covariance_matrix)
```

### Prior of classes

```{r, message=FALSE , error=FALSE,echo=FALSE}

setosa_prior <- nrow(setosa)/nrow(data)
setosa_prior
versicolor_prior <- nrow(versicolor)/nrow(data)
versicolor_prior
virginica_prior <- nrow(virginica)/nrow(data)
virginica_prior
```


### Pooled covariance

```{r , message=FALSE , error=FALSE}
pooled_matrix <- (virginica_covariance*nrow(virginica)+(setosa_covariance*nrow(setosa))+versicolor_covariance*nrow(versicolor))/150
pooled_matrix
```

### Underlying probabbilistic model

$$x|y=C_i,\mu_i,\sum ~ N(\mu_i,\sum)$$
$$y|\pi =  Multinomial(\pi_i.....\pi_k)$$


### Discriminant function

```{r  , message=FALSE , error=FALSE}

disc_function <- function(class_data,class_mean,class_prior,class_covariance)
{
  (as.matrix(class_data[1:2]) %*% solve(class_covariance) %*% class_mean) -     
  rep((0.5*(class_mean[1:2]%*%solve(class_covariance)%*%class_mean)),nrow(class_data))+rep(log(class_prior),nrow(class_data))
  
}

setosa_disc <- disc_function(data,setosa_mean,setosa_prior,pooled_matrix)
versicolor_disc <-  disc_function(data,versicolor_mean,versicolor_prior,pooled_matrix)
virginica_disc <- disc_function(data,virginica_mean,virginica_prior,pooled_matrix)


```


### Decision boundaries 

$$Log(\pi_l/\pi_k)-1/2 (\mu_k+\mu_l)^T\sum(\mu_k-\mu_l)+x^T\sum^{-1}(\mu_k-\mu_l)$$

```{r,echo = FALSE , message=FALSE , error=FALSE,results='hide',warning=FALSE}
decision_boundry <- function(pi_k,pi_l,mu_1,mu_2,x){
  rep(log(pi_k/pi_l))-rep((0.5*t((mu_1+mu_2))%*%mu_1-mu_2)) + (as.matrix(x[1:2])%*%(mu_1-mu_2))
}

decision_boundry(setosa_prior,versicolor_prior,setosa_mean,versicolor_mean,data)
decision_boundry(setosa_prior,virginica_prior,setosa_mean,virginica_mean,data)
decision_boundry(versicolor_prior,virginica_prior,versicolor_mean,virginica_mean,data)

```



### Predicting using dicriminant function

```{r echo = FALSE , message=FALSE , error=FALSE,results='hide',warning=FALSE}
dis_val <- cbind(setosa_disc,versicolor_disc,virginica_disc)
dis_val <- as.data.frame(dis_val)

names(dis_val) <- c("Setosa" , "versicolor" , "virginica")

values <- colnames(dis_val)[max.col(dis_val,ties.method = "random")]
dis_val$values <- values


ggplot(data,aes(x=Sepal.Length , y = Sepal.Width,color = dis_val$values))+
  geom_point()

```


### Prediction using LDA

```{r,echo = FALSE , message=FALSE , error=FALSE}


model_lda <- lda(Species ~ Sepal.Length+Sepal.Width , data = data)

lda_prediction <- predict(model_lda , data[1:2])

lda_cfm <- table(lda_prediction$class,data$Species)


ggplot(data ,aes(x=Sepal.Width,y=Sepal.Length , color  = lda_prediction$class))+
  geom_point()


lda_miscl_error <- 1-sum(diag(lda_cfm))/sum(lda_cfm)


```

## Confusion matrix using LDA function

```{r,echo = FALSE , message=FALSE , error=FALSE}
lda_cfm
```

## Misclassification error using LDA function

```{r,echo=FALSE}
lda_miscl_error
```


### Generatin new data with discriminant function


```{r,echo = FALSE , message=FALSE , error=FALSE,results='hide',warning=FALSE}
library(mvtnorm)
  

n_s <- rmvnorm(50, mean = setosa_mean, sigma = setosa_covariance )

n_ve <- rmvnorm(50, mean = versicolor_mean, sigma = versicolor_covariance )

n_vi <- rmvnorm(50, mean = virginica_mean, sigma = virginica_covariance )

  n_all<-rbind(n_s,n_ve,n_vi)
  
  n_all<-as.data.frame(n_all)
  
  n_all$values <- values
  
  
  sample_data <-  n_all[sample(nrow(n_all), 150), ]
  
  
  names(n_all)<-c("sepal.length","sepal.width","species")
  

  ggplot(data = n_all, aes(x = sepal.length,y = sepal.width,color = values))+
           geom_point()

```

### Prediction using logistic regression

```{r,echo = FALSE , message=FALSE , error=FALSE,warning=FALSE}

library(nnet)

tr_data <- data
tr_data$Species <- relevel(tr_data$Species, ref = 'setosa')

mlm_model <- multinom(Species ~ Sepal.Length + Sepal.Width , data  = tr_data)


pred <- predict(mlm_model , tr_data)

confusion_matrix <- table(predict(mlm_model),tr_data$Species)



misclassification_error <- 1-sum(diag(confusion_matrix))/sum(confusion_matrix)

```

#### Confusion matrix

```{r echo = FALSE , message=FALSE , error=FALSE}
confusion_matrix
```

#### Misclassification error

```{r echo = FALSE , message=FALSE , error=FALSE}
misclassification_error

```




