---
title: "Neural Network"
author: "Varshith konda"
date: "20/6/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


# NEURAL NETWORKS

Objective of the following assignment is to  use neuralnet package from CRAN to implement neural network.

Our data is 500 uniformly generated random points in the interval [0,10]



```{r , echo = FALSE , warning=FALSE ,error=FALSE}

library(neuralnet)
library(ggplot2)


set.seed(1234567890)
Var <- runif(500, 0, 10)
mydata <- data.frame(Var, Sin=sin(Var))
tr <- mydata[1:25,] # Training
te <- mydata[26:500,] # Test

# Random initialization of the weights in the interval [-1, 1]

winit <- c(-1,1)

```


## Training model on training data



### Error,threshold and steps data of model during training

```{r, echo = FALSE , warning=FALSE ,error=FALSE}


# Training neural network with training data


nn <- neuralnet(Sin~Var , 
                data = tr , 
                hidden = 5 ,
                startweights = winit,
                lifesign = "minimal",
                linear.output = T,
                rep = 10 )


```

Here sin values of the points are response variables and points are used as feature for training.

After generating 500 sample points , we are using just 5 percent of the data to train the neural network and 95 percent of the data for testing the accuracy of the network.

We are not using any activation function during model training as we our main motive is not classification , this is the reason "Linear.output" is  TRUE.

Five hidden layers and 10 repetitions were carried out during training process , default weights of the network are  set to  NULL but we are randomly initializing weights with [-1,1].

Repetition with lowest error rate and lowest number of steps is used for predicting test values.


## Plotting true training values and predicted values from model.



```{r, echo = FALSE , warning=FALSE ,error=FALSE}
# predicting the training data with the model
pr <- compute(nn,tr,rep = 5)

# plotting the predicited values and true values
ggplot()+
  geom_point(aes(x=tr$Var,y=tr$Sin,color= "True-vlues"))+
  geom_point(aes(x=tr$Var,y=pr$net.result,color= "Predicted-vlues"))

```

Network performed well for training data with slight margin of errors.

## Predicting test data using trained  model





```{r, echo = FALSE , warning=FALSE ,error=FALSE}

# Predicting test data with model
pr_test <- compute(nn,te,rep = 5)

# plotting true values and predicted values
ggplot()+
  geom_point(aes(x=te$Var,y=te$Sin,color= "True-data"))+
  geom_point(aes(x=te$Var,y=pr_test$net.result , color = "Predicted-values"))+
  labs(x  = "Variable" , y = "Sin ")
```



Neural net performed moderately with minimal errors , model underperformed while predicting values from 5 to 7.5 


## Predicting from newly generated data  


```{r, echo = FALSE , warning=FALSE ,error=FALSE}
set.seed(1234567890)

# Generating new values from 0  - 20
Var <- runif(500,0,20)

# Appending to data frame
twenty_500 <- data.frame(Var ,Sin =  sin(Var))

# Predicting new data with model
pred_new <- compute(nn,twenty_500,rep=5)

# plotting new test values with predicted values
ggplot()+
  geom_point(data = twenty_500 , aes(x = Var , y = Sin,color = "True-values" ))+
  geom_point(aes(x = twenty_500$Var , y = pred_new$net.result, color = "Predicted-values" ))+
  labs(x  = "Variable" , y = "Sin ")


```



New data was sampled from uniformly distribution in the interval of 0 to 20.

While comparing the predicted values from model with respect to true values, model accurately predicted the values until midway.


Our sample size for  training data is very low which can cause "under fitting" , model cannot understand the coherence between features and response variable. 

Due to the under fitting of the data error rate of the model can be higher.


## Predicting variable from sin values 


```{r, echo = FALSE , warning=FALSE ,error=FALSE}
set.seed(12345678)
# generating new data
 

# appending to data frame
new_Var <- runif(500, 0, 10)
new_data <- data.frame(new_Var, new_Sine=sin(new_Var))

new_nn <- neuralnet(new_Var~new_Sine , 
                data = new_data , 
                hidden = 5 ,
                startweights = winit,
                lifesign = "minimal",
                rep = 10,
                threshold = 0.03,
                linear.output = T)

# predictintg value
new_nn_pred <- compute(new_nn , new_data,rep = 3)

# plotting true vales with predicted values
ggplot()+
  geom_point(data = new_data ,aes(x = new_Sine , y = new_Var , color = "True-values"))+
  geom_point(aes(x = new_data$new_Sine , y= new_nn_pred$net.result,color = "Predicted-values"))+
  labs(x  =  "Sin " , y ="Variable")
```


The algorithm was not converging within the step max range, so the threshold was increased to 0.03 for converging the algorithm within the step max range.

The value on y can have many possibilities from one sin function, which can lead to an improper working model.


